---
title: "Pet Store Sales"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_float: yes
    number_sections: no
---

This is data from a pet store's in-person and online sales.

Source: [Pet Store](https://github.com/alanhlambda/dataset/raw/main/pet%20store.xlsx)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)

inu=import('https://github.com/alanhlambda/dataset/raw/main/pet%20store.xlsx')

sum(is.na(inu))
```

No missing values.

## Customer Retention

```{r}
sum(duplicated(inu$cust_id))/sum(!duplicated(inu$cust_id))*100

```

79.95% customer retention rate.

## Customer Age

```{r}
inu |> 
  ggplot(aes(x=cust_age))+
  geom_bar(color='darkred')+
  labs(y='Total Count',x='Customer Ages')

```

Mid-20s are the biggest group of customers.

## Customer Location

```{r}
inu %>% 
  rename(Customer_State=cust_state) %>% 
  tabyl(Customer_State) %>%
  arrange(desc(n))

inu %>%
  ggplot(aes(y=cust_state))+
  theme_bw()+
  geom_bar(color='darkblue')+
  labs(y='States',x='Total Count')
```

NY, PA, and NJ are the top three states where most of the customers order from.

## Product Sales

```{r}
inu %>% 
  rename(Product_Category=prod_category) %>% 
  tabyl(Product_Category,prod_animal_type) %>%
  adorn_totals() %>% 
  adorn_percentages('all') %>%
  adorn_pct_formatting(digits = 1,affix_sign = T) %>%
  adorn_ns(position='front') %>% 
  arrange(desc(cat))
```

```{r}
inu %>% 
  ggplot(aes(x=prod_category,fill=prod_animal_type))+
  geom_bar()+
  theme_bw()+
  labs(title='Number of Items Sold',x='Product Category',y='Total Count',fill=' ')+
  theme(plot.title = element_text(hjust=.5))

```

Cat products, especially cat treats, are more in demand than dog products.

```{r,warning=F,message=F}
inu %>% 
  group_by('Product Title'=prod_title,'Product Category'=prod_category,'Animal Type'=prod_animal_type) %>%   summarise(Total_Sales=sum(total_sales)) %>% 
  arrange(desc(Total_Sales))

```

Reddy Beddy and Cat Cave are the most popular items sold.

```{r}
sessionInfo()
```
